My title\thanks{Code and data are available at: \url{https://github.com/RohanAlexander/starter_folder}.}

### 2024-11-14 paper/paper.qmd
Added fig-label

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

@@

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

@@

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

@@

```{r}
#### Read data ####
# Load the cleaned dataset
analysis_data <- read_csv("C:/Users/eliza/Downloads/hammer-3-compressed/cleaned_price_changes.csv", show_col_types = FALSE)

@@

Figure 1: Average Current Price by Vendor. This bar plot shows the average price of products sold by each grocery vendor in Canadian dollars (CAD). Variations in pricing suggest differences in market positioning or pricing strategies among vendors, with some retailers consistently offering higher or lower prices across product categories.

@@

```{r}
# Average price difference by vendor
price_diff_plot <- analysis_data %>%
  group_by(vendor_name) %>%
  summarize(avg_price_difference = mean(price_difference, na.rm = TRUE)) %>%
  ggplot(aes(x = reorder(vendor_name, avg_price_difference), y = avg_price_difference)) +
  geom_bar(stat = "identity", fill = "coral") +
  coord_flip() +
  labs(
    title = "Average Price Difference by Vendor",
    x = "Vendor",
    y = "Average Price Difference (CAD)"
  ) +
  theme_minimal()
price_diff_plot

@@

Figure 2: Average Price Difference by Vendor. This bar plot shows the average price change per vendor, highlighting retailers with the most significant pricing adjustments. Positive differences suggest recent price increases, while negative differences indicate reductions, reflecting possible discounting or competitive pricing strategies.

@@

```{r}
# Scatter plot of previous price vs. current price
price_relationship_plot <- ggplot(analysis_data, aes(x = price_old, y = price_current)) +
  geom_point(alpha = 0.4, color = "blue") +
  geom_smooth(method = "lm", color = "red") +
  labs(
    title = "Relationship Between Previous and Current Prices",
    x = "Previous Price (CAD)",
    y = "Current Price (CAD)"
  ) +
  theme_minimal()
price_relationship_plot

```
Figure 3: Relationship Between Previous and Current Prices. This scatter plot with a linear regression line shows the correlation between previous and current prices. A positive relationship would indicate that higher previous prices predict higher current prices, implying potential price stability or gradual price changes in the grocery market.

@@

```{r}
# Calculate mean and standard deviation of price_difference
mean_diff <- mean(analysis_data$price_difference, na.rm = TRUE)
sd_diff <- sd(analysis_data$price_difference, na.rm = TRUE)

@@

Figure 4: Distribution of Price Differences (Filtered). This histogram illustrates the frequency of price changes across products, with extreme values beyond three standard deviations removed for clarity. The concentration around zero suggests that small price adjustments are most common, while the presence of both negative and positive values indicates instances of both price increases and decreases. This distribution provides insight into the stability and variability of pricing across vendors.
## Model Evaluation

To evaluate our model’s performance, we assess the predictive accuracy of the Bayesian linear regression model on the test data. The model’s Mean Absolute Error (MAE) and Root Mean Squared Error (RMSE) provide insights into the model’s ability to predict `price_current` based on `price_old`, `vendor_name`, and `price_difference`.

```{r}

#### Read Data ####
# Load only a sample of the data to reduce memory usage (e.g., 10% of the dataset)
set.seed(853)

analysis_data <- fread("C:/Users/eliza/Downloads/hammer-3-compressed/cleaned_price_changes.csv")
analysis_data <- analysis_data[sample(.N, size = 0.1 * .N)]

# Split data into training and testing sets
train_indices <- sample(seq_len(nrow(analysis_data)), size = 0.8 * nrow(analysis_data))
analysis_data_train <- analysis_data[train_indices]
analysis_data_test <- analysis_data[-train_indices]

#### Model Data ####
# Define a Bayesian regression model to predict `price_current` with fewer iterations and lighter priors

model <- stan_glm(
  formula = price_current ~ price_old + as.factor(vendor_name) + price_difference,
  data = analysis_data_train,
  family = gaussian(),
  prior = normal(location = 0, scale = 5, autoscale = TRUE),      # Slightly broader prior
  prior_intercept = normal(location = 0, scale = 5, autoscale = TRUE),
  prior_aux = exponential(rate = 0.5, autoscale = TRUE),          # Less informative prior
  iter = 500,                                                     # Reduce iterations
  chains = 2,                                                     # Fewer chains for faster computation
  seed = 853
)

#### Predict in Chunks ####
# Define chunk size based on available memory
chunk_size <- 20000  # Adjusted smaller chunk size for efficiency
num_chunks <- ceiling(nrow(analysis_data_test) / chunk_size)

# Initialize a vector to store predictions
predictions <- numeric(nrow(analysis_data_test))

# Loop through each chunk to generate predictions
for (i in seq_len(num_chunks)) {
  start <- ((i - 1) * chunk_size) + 1
  end <- min(i * chunk_size, nrow(analysis_data_test))

  # Generate predictions for the current chunk
  predictions[start:end] <- predict(model, newdata = analysis_data_test[start:end, ])
}

#### Model Evaluation Metrics ####
# Calculate Mean Absolute Error (MAE) and Root Mean Squared Error (RMSE)
mae <- mean(abs(predictions - analysis_data_test$price_current))
rmse <- sqrt(mean((predictions - analysis_data_test$price_current)^2))

# Store results in a data frame
evaluation_metrics <- data.table(
  Metric = c("Mean Absolute Error (MAE)", "Root Mean Squared Error (RMSE)"),
  Value = c(round(mae, 2), round(rmse, 2))
)

# Display the results table
print(evaluation_metrics)

Table 1: Model Evaluation Metrics. The table displays the MAE and RMSE for the model, reflecting the average and squared prediction errors, respectively. Lower values indicate better predictive accuracy, suggesting that the model effectively captures the relationship between previous prices, vendor identities, and price differences.

These results demonstrate that vendor identity and previous prices are significant predictors of current prices in the Canadian grocery sector, with implications for understanding competitive strategies and consumer price sensitivity in the market.

>>>>

@@

@@

@@

```{r}
#| label: fig-avg-curr-price
#| fig-cap: Average Current Price by Vendor. This bar plot shows the average price of products sold by each grocery vendor in Canadian dollars (CAD). Variations in pricing suggest differences in market positioning or pricing strategies among vendors, with some retailers consistently offering higher or lower prices across product categories.
#| echo: false
#| warning: false
#| message: false

#### Read data ####
# Load the cleaned dataset
analysis_data <- read_csv("C:/Users/eliza/Downloads/hammer-3-compressed/cleaned_price_changes.csv", show_col_types = FALSE)

@@

@@

```{r}
#| label: fig-avg-price-diff
#| fig-cap: Average Price Difference by Vendor. This bar plot shows the average price change per vendor, highlighting retailers with the most significant pricing adjustments. Positive differences suggest recent price increases, while negative differences indicate reductions, reflecting possible discounting or competitive pricing strategies.
#| echo: false
#| warning: false
#| message: false

# Average price difference by vendor
price_diff_plot <- analysis_data %>%
  group_by(vendor_name) %>%
  summarize(avg_price_difference = mean(price_difference, na.rm = TRUE)) %>%
  ggplot(aes(x = reorder(vendor_name, avg_price_difference), y = avg_price_difference)) +
  geom_bar(stat = "identity", fill = "coral") +
  coord_flip() +
  labs(
    title = "Average Price Difference by Vendor",
    x = "Vendor",
    y = "Average Price Difference (CAD)"
  ) +
  theme_minimal()
price_diff_plot

@@

@@

```{r}
#| label: fig-past-curr
#| fig-cap: Relationship Between Previous and Current Prices. This scatter plot with a linear regression line shows the correlation between previous and current prices. A positive relationship would indicate that higher previous prices predict higher current prices, implying potential price stability or gradual price changes in the grocery market.
#| echo: false
#| warning: false
#| message: false

# Scatter plot of previous price vs. current price
price_relationship_plot <- ggplot(analysis_data, aes(x = price_old, y = price_current)) +
  geom_point(alpha = 0.4, color = "blue") +
  geom_smooth(method = "lm", color = "red") +
  labs(
    title = "Relationship Between Previous and Current Prices",
    x = "Previous Price (CAD)",
    y = "Current Price (CAD)"
  ) +
  theme_minimal()
price_relationship_plot

@@

```{r}
#| label: fig-dist-price
#| fig-cap: Relationship Between Previous and Current Prices. This scatter plot with a linear regression line shows the correlation between previous and current prices. A positive relationship would indicate that higher previous prices predict higher current prices, implying potential price stability or gradual price changes in the grocery market.
#| echo: false
#| warning: false
#| message: false

# Calculate mean and standard deviation of price_difference
mean_diff <- mean(analysis_data$price_difference, na.rm = TRUE)
sd_diff <- sd(analysis_data$price_difference, na.rm = TRUE)

@@

### 2024-11-14 paper/paper.tex
Added fig-label

Figure 4: Distribution of Price Differences (Filtered). This histogram
illustrates the frequency of price changes across products, with extreme
values beyond three standard deviations removed for clarity. The
concentration around zero suggests that small price adjustments are most
common, while the presence of both negative and positive values
indicates instances of both price increases and decreases. This
distribution provides insight into the stability and variability of
pricing across vendors. \#\# Model Evaluation

To evaluate our model's performance, we assess the predictive accuracy
of the Bayesian linear regression model on the test data. The model's
Mean Absolute Error (MAE) and Root Mean Squared Error (RMSE) provide
insights into the model's ability to predict \texttt{price\_current}
based on \texttt{price\_old}, \texttt{vendor\_name}, and
\texttt{price\_difference}.

\begin{Shaded}
\begin{Highlighting}[]
\DocumentationTok{\#\#\#\# Read Data \#\#\#\#}
\CommentTok{\# Load only a sample of the data to reduce memory usage (e.g., 10\% of the dataset)}
\FunctionTok{set.seed}\NormalTok{(}\DecValTok{853}\NormalTok{)}

\NormalTok{analysis\_data }\OtherTok{\textless{}{-}} \FunctionTok{fread}\NormalTok{(}\StringTok{"C:/Users/eliza/Downloads/hammer{-}3{-}compressed/cleaned\_price\_changes.csv"}\NormalTok{)}
\NormalTok{analysis\_data }\OtherTok{\textless{}{-}}\NormalTok{ analysis\_data[}\FunctionTok{sample}\NormalTok{(.N, }\AttributeTok{size =} \FloatTok{0.1} \SpecialCharTok{*}\NormalTok{ .N)]}

\CommentTok{\# Split data into training and testing sets}
\NormalTok{train\_indices }\OtherTok{\textless{}{-}} \FunctionTok{sample}\NormalTok{(}\FunctionTok{seq\_len}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(analysis\_data)), }\AttributeTok{size =} \FloatTok{0.8} \SpecialCharTok{*} \FunctionTok{nrow}\NormalTok{(analysis\_data))}
\NormalTok{analysis\_data\_train }\OtherTok{\textless{}{-}}\NormalTok{ analysis\_data[train\_indices]}
\NormalTok{analysis\_data\_test }\OtherTok{\textless{}{-}}\NormalTok{ analysis\_data[}\SpecialCharTok{{-}}\NormalTok{train\_indices]}

\DocumentationTok{\#\#\#\# Model Data \#\#\#\#}
\CommentTok{\# Define a Bayesian regression model to predict \textasciigrave{}price\_current\textasciigrave{} with fewer iterations and lighter priors}
\FunctionTok{library}\NormalTok{(rstanarm)}
\NormalTok{model }\OtherTok{\textless{}{-}} \FunctionTok{stan\_glm}\NormalTok{(}
  \AttributeTok{formula =}\NormalTok{ price\_current }\SpecialCharTok{\textasciitilde{}}\NormalTok{ price\_old }\SpecialCharTok{+} \FunctionTok{as.factor}\NormalTok{(vendor\_name) }\SpecialCharTok{+}\NormalTok{ price\_difference,}
  \AttributeTok{data =}\NormalTok{ analysis\_data\_train,}
  \AttributeTok{family =} \FunctionTok{gaussian}\NormalTok{(),}
  \AttributeTok{prior =} \FunctionTok{normal}\NormalTok{(}\AttributeTok{location =} \DecValTok{0}\NormalTok{, }\AttributeTok{scale =} \DecValTok{5}\NormalTok{, }\AttributeTok{autoscale =} \ConstantTok{TRUE}\NormalTok{),      }\CommentTok{\# Slightly broader prior}
  \AttributeTok{prior\_intercept =} \FunctionTok{normal}\NormalTok{(}\AttributeTok{location =} \DecValTok{0}\NormalTok{, }\AttributeTok{scale =} \DecValTok{5}\NormalTok{, }\AttributeTok{autoscale =} \ConstantTok{TRUE}\NormalTok{),}
  \AttributeTok{prior\_aux =} \FunctionTok{exponential}\NormalTok{(}\AttributeTok{rate =} \FloatTok{0.5}\NormalTok{, }\AttributeTok{autoscale =} \ConstantTok{TRUE}\NormalTok{),          }\CommentTok{\# Less informative prior}
  \AttributeTok{iter =} \DecValTok{500}\NormalTok{,                                                     }\CommentTok{\# Reduce iterations}
  \AttributeTok{chains =} \DecValTok{2}\NormalTok{,                                                     }\CommentTok{\# Fewer chains for faster computation}
  \AttributeTok{seed =} \DecValTok{853}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}

SAMPLING FOR MODEL 'continuous' NOW (CHAIN 1).
Chain 1: 
Chain 1: Gradient evaluation took 0.000105 seconds
Chain 1: 1000 transitions using 10 leapfrog steps per transition would take 1.05 seconds.
Chain 1: Adjust your expectations accordingly!
Chain 1: 
Chain 1: 
Chain 1: Iteration:   1 / 500 [  0%]  (Warmup)
Chain 1: Iteration:  50 / 500 [ 10%]  (Warmup)
Chain 1: Iteration: 100 / 500 [ 20%]  (Warmup)
Chain 1: Iteration: 150 / 500 [ 30%]  (Warmup)
Chain 1: Iteration: 200 / 500 [ 40%]  (Warmup)
Chain 1: Iteration: 250 / 500 [ 50%]  (Warmup)
Chain 1: Iteration: 251 / 500 [ 50%]  (Sampling)
Chain 1: Iteration: 300 / 500 [ 60%]  (Sampling)
Chain 1: Iteration: 350 / 500 [ 70%]  (Sampling)
Chain 1: Iteration: 400 / 500 [ 80%]  (Sampling)
Chain 1: Iteration: 450 / 500 [ 90%]  (Sampling)
Chain 1: Iteration: 500 / 500 [100%]  (Sampling)
Chain 1: 
Chain 1:  Elapsed Time: 0.056 seconds (Warm-up)
Chain 1:                3.902 seconds (Sampling)
Chain 1:                3.958 seconds (Total)
Chain 1:

SAMPLING FOR MODEL 'continuous' NOW (CHAIN 2).
Chain 2: 
Chain 2: Gradient evaluation took 2.5e-05 seconds
Chain 2: 1000 transitions using 10 leapfrog steps per transition would take 0.25 seconds.
Chain 2: Adjust your expectations accordingly!
Chain 2: 
Chain 2: 
Chain 2: Iteration:   1 / 500 [  0%]  (Warmup)
Chain 2: Iteration:  50 / 500 [ 10%]  (Warmup)
Chain 2: Iteration: 100 / 500 [ 20%]  (Warmup)
Chain 2: Iteration: 150 / 500 [ 30%]  (Warmup)
Chain 2: Iteration: 200 / 500 [ 40%]  (Warmup)
Chain 2: Iteration: 250 / 500 [ 50%]  (Warmup)
Chain 2: Iteration: 251 / 500 [ 50%]  (Sampling)
Chain 2: Iteration: 300 / 500 [ 60%]  (Sampling)
Chain 2: Iteration: 350 / 500 [ 70%]  (Sampling)
Chain 2: Iteration: 400 / 500 [ 80%]  (Sampling)
Chain 2: Iteration: 450 / 500 [ 90%]  (Sampling)
Chain 2: Iteration: 500 / 500 [100%]  (Sampling)
Chain 2: 
Chain 2:  Elapsed Time: 0.036 seconds (Warm-up)
Chain 2:                3.936 seconds (Sampling)
Chain 2:                3.972 seconds (Total)
Chain 2: 
\end{verbatim}

\begin{verbatim}
Warning: There were 1 chains where the estimated Bayesian Fraction of Missing Information was low. See
https://mc-stan.org/misc/warnings.html#bfmi-low
\end{verbatim}

\begin{verbatim}
Warning: Examine the pairs() plot to diagnose sampling problems
\end{verbatim}

\begin{verbatim}
Warning: The largest R-hat is 1.6, indicating chains have not mixed.
Running the chains for more iterations may help. See
https://mc-stan.org/misc/warnings.html#r-hat
\end{verbatim}

\begin{verbatim}
Warning: Bulk Effective Samples Size (ESS) is too low, indicating posterior means and medians may be unreliable.
Running the chains for more iterations may help. See
https://mc-stan.org/misc/warnings.html#bulk-ess
\end{verbatim}

\begin{verbatim}
Warning: Tail Effective Samples Size (ESS) is too low, indicating posterior variances and tail quantiles may be unreliable.
Running the chains for more iterations may help. See
https://mc-stan.org/misc/warnings.html#tail-ess
\end{verbatim}

\begin{verbatim}
Warning: Markov chains did not converge! Do not analyze results!
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\DocumentationTok{\#\#\#\# Predict in Chunks \#\#\#\#}
\CommentTok{\# Define chunk size based on available memory}
\NormalTok{chunk\_size }\OtherTok{\textless{}{-}} \DecValTok{20000}  \CommentTok{\# Adjusted smaller chunk size for efficiency}
\NormalTok{num\_chunks }\OtherTok{\textless{}{-}} \FunctionTok{ceiling}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(analysis\_data\_test) }\SpecialCharTok{/}\NormalTok{ chunk\_size)}

\CommentTok{\# Initialize a vector to store predictions}
\NormalTok{predictions }\OtherTok{\textless{}{-}} \FunctionTok{numeric}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(analysis\_data\_test))}

\CommentTok{\# Loop through each chunk to generate predictions}
\ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \FunctionTok{seq\_len}\NormalTok{(num\_chunks)) \{}
\NormalTok{  start }\OtherTok{\textless{}{-}}\NormalTok{ ((i }\SpecialCharTok{{-}} \DecValTok{1}\NormalTok{) }\SpecialCharTok{*}\NormalTok{ chunk\_size) }\SpecialCharTok{+} \DecValTok{1}
\NormalTok{  end }\OtherTok{\textless{}{-}} \FunctionTok{min}\NormalTok{(i }\SpecialCharTok{*}\NormalTok{ chunk\_size, }\FunctionTok{nrow}\NormalTok{(analysis\_data\_test))}

  \CommentTok{\# Generate predictions for the current chunk}
\NormalTok{  predictions[start}\SpecialCharTok{:}\NormalTok{end] }\OtherTok{\textless{}{-}} \FunctionTok{predict}\NormalTok{(model, }\AttributeTok{newdata =}\NormalTok{ analysis\_data\_test[start}\SpecialCharTok{:}\NormalTok{end, ])}
\NormalTok{\}}

\DocumentationTok{\#\#\#\# Model Evaluation Metrics \#\#\#\#}
\CommentTok{\# Calculate Mean Absolute Error (MAE) and Root Mean Squared Error (RMSE)}
\NormalTok{mae }\OtherTok{\textless{}{-}} \FunctionTok{mean}\NormalTok{(}\FunctionTok{abs}\NormalTok{(predictions }\SpecialCharTok{{-}}\NormalTok{ analysis\_data\_test}\SpecialCharTok{$}\NormalTok{price\_current))}
\NormalTok{rmse }\OtherTok{\textless{}{-}} \FunctionTok{sqrt}\NormalTok{(}\FunctionTok{mean}\NormalTok{((predictions }\SpecialCharTok{{-}}\NormalTok{ analysis\_data\_test}\SpecialCharTok{$}\NormalTok{price\_current)}\SpecialCharTok{\^{}}\DecValTok{2}\NormalTok{))}

\CommentTok{\# Store results in a data frame}
\NormalTok{evaluation\_metrics }\OtherTok{\textless{}{-}} \FunctionTok{data.table}\NormalTok{(}
  \AttributeTok{Metric =} \FunctionTok{c}\NormalTok{(}\StringTok{"Mean Absolute Error (MAE)"}\NormalTok{, }\StringTok{"Root Mean Squared Error (RMSE)"}\NormalTok{),}
  \AttributeTok{Value =} \FunctionTok{c}\NormalTok{(}\FunctionTok{round}\NormalTok{(mae, }\DecValTok{2}\NormalTok{), }\FunctionTok{round}\NormalTok{(rmse, }\DecValTok{2}\NormalTok{))}
\NormalTok{)}

\CommentTok{\# Display the results table}
\FunctionTok{print}\NormalTok{(evaluation\_metrics)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
                           Metric Value
                           <char> <num>
1:      Mean Absolute Error (MAE)  6.74
2: Root Mean Squared Error (RMSE) 17.98
\end{verbatim}

Table 1: Model Evaluation Metrics. The table displays the MAE and RMSE
for the model, reflecting the average and squared prediction errors,
respectively. Lower values indicate better predictive accuracy,
suggesting that the model effectively captures the relationship between
previous prices, vendor identities, and price differences.

These results demonstrate that vendor identity and previous prices are
significant predictors of current prices in the Canadian grocery sector,
with implications for understanding competitive strategies and consumer
price sensitivity in the market.

>>>>

Figure 4: Distribution of Price Differences (Filtered). This histogram
illustrates the frequency of price changes across products, with extreme
values beyond three standard deviations removed for clarity. The
concentration around zero suggests that small price adjustments are most
common, while the presence of both negative and positive values
indicates instances of both price increases and decreases. This
distribution provides insight into the stability and variability of
pricing across vendors.
